### 2022-11-07 ic/aistats2021/paper.tex
Updates from Overleaf

We next consider a Gaussian likelihood with a Gaussian mean prior, a
conjugate model with closed-form posterior given by:
\begin{align}
  x                                 \sim \mathcal{N}(0, \sigma_x),
                                    & \qquad y \sim \mathcal{N}(x, \sigma_y) \label{eq:conj-normal} \\
  \Pr[x \mid y, \sigma_x, \sigma_y] & \sim \mathcal{N}\left(
  \frac{\sigma_y^{-2}}{\sigma_x^{-2} + \sigma_{y}^{-2}} y,
  \frac{1}{\sigma_x^{-2} + \sigma_{y}^{-2}}
  \right) \nonumber
\end{align}
\begin{figure}
  \centering
  \includegraphics[width=0.8\linewidth]{Figures/normal_normal_mean_type1.pdf}
  \caption{In a conjugate normal-normal model, LIC's proposal distribution
    mean (dashed orange line) closely follows the closed-form posterior mean
    (blue solid line) across a wide range of observed values $y$.}\label{fig:normal-normal}
\end{figure}

@@

Consider a Bayesian logistic regression model over $d$ covariates
with prior
$\vec{\beta} \sim \cN_{d+1}(\vec{0}_{d+1}, \diag(10, 2.5\vec{1}_{d}))$
and likelihood $y_i \mid \vx_i \simiid \text{Bernoulli}(\sigma(\vec{\beta}^\top \vx_i))$
where $\sigma(t) = (1 + e^{-t})^{-1}$ is the logistic function.
This model is appropriate in a wide range of classification problems where prior
knowledge about the regression coefficients $\vec{\beta}$ are available.

@@

\begin{figure*}
  \centering
  \caption{
    Results drawing 100 samples across 10 chains (after 1,000 burn-in / adaptation / compilation samples)
    on two Bayesian inference problems.
    The top row of each subfigure aggregates over the 10 chains the (left to
    right) compilation time (time to instantiate the Python class, includes
    neural network training for LIC and Stan's C\texttt{++} codegen and compilation
    for NUTS), inference time (time from invoking \texttt{infer()} to when
    sampling terminates), and the predictive log-likelihood (PLL) on held-out
    test data.
    The bottom row aggregates over both the 10 chains as well as over all
    latent variables the (left to right) estimated expected sample size (ESS,
    higher is better, \cite{geyer2011introduction}) and the rank normalized
    $\widehat{R}$ diagnostic (Rhat, closer to $1$ is better, \cite{vehtari2020rank}).
  }\label{fig:blr_nschools}
  \begin{subfigure}[b]{\textwidth}
    \centering
    \caption{Bayesian logistic regression ($2000$ rows, $10$ features).
      Both LIC and Stan (NUTS) amortize the upfront compilation cost with accelerated inference times.
      LIC achieves comparable PLL to NUTS \cite{hoffman2014no} and other methods,
      and yields ESS comparable to NUTS and higher than any other method.
      The $\widehat{R}$ of LIC is close to $1$ (similar to NUTS and lower than
      all other methods).
    }\label{fig:blr}
    \includegraphics[width=\linewidth]{Figures/blr_pll_type1.pdf}
    \includegraphics[width=0.6\linewidth]{Figures/blr_ess_rhat_type1.pdf}
  \end{subfigure}
  \begin{subfigure}[b]{\textwidth}
    \centering
    \caption{n-schools (1000 schools, 8 states, 5 districts, 5 types).
      Again, increased compilation times are offset by accelerated inference times.
      In this case, LIC achieves comparable PLL to NUTS while simultaneously producing
      higher ESS and lower $\widehat{R}$, suggesting that the resulting posterior samples
      are less autocorrelated and provide a more accurate posterior approximation.
    }\label{fig:nschools}
    \includegraphics[width=\linewidth]{Figures/nschools_pll_type1.pdf}
    \includegraphics[width=0.66\linewidth]{Figures/nschools_ess_rhat_type1.pdf}
  \end{subfigure}
\end{figure*}

@@

% TODO: include in camera ready
\subsubsection*{Acknowledgements}
Feynman Liang is supported by a NPSC Graduate Fellowship.
This work was completed during an internship at Facebook.

>>>>

We next consider a Gaussian likelihood with a Gaussian mean prior, a
conjugate model with closed-form posterior given by:
\begin{align}
  x                                 \sim \mathcal{N}(0, \sigma_x),
                                    & \qquad y \sim \mathcal{N}(x, \sigma_y) \label{eq:conj-normal} \\
  \Pr[x \mid y, \sigma_x, \sigma_y] & \sim \mathcal{N}\left(
  \frac{\sigma_y^{-2}}{\sigma_x^{-2} + \sigma_{y}^{-2}} y,
  \frac{1}{\sigma_x^{-2} + \sigma_{y}^{-2}}
  \right) \nonumber
\end{align}
\begin{figure}
  \centering
  \includegraphics[width=0.6\linewidth]{Figures/normal_normal_mean_type1.pdf}
  \caption{In a conjugate normal-normal model, LIC's proposal distribution
    mean (dashed orange line) closely follows the closed-form posterior mean
    (blue solid line) across a wide range of observed values $y$.}\label{fig:normal-normal}
\end{figure}

@@

Consider a Bayesian logistic regression model over $d$ covariates
with prior distribution 
$\vec{\beta} \sim \cN_{d+1}(\vec{0}_{d+1}, \diag(10, 2.5\vec{1}_{d}))$
and likelihood $y_i \mid \vx_i \simiid \text{Bernoulli}(\sigma(\vec{\beta}^\top \vx_i))$
where $\sigma(t) = (1 + e^{-t})^{-1}$ is the logistic function.
This model is appropriate in a wide range of classification problems where prior
knowledge about the regression coefficients $\vec{\beta}$ are available.

@@

\begin{figure*}
  \caption{
    Results of MCMC on two Bayesian inference tasks where we compare
    the compilation time (neural network training for LIC and Stan's C\texttt{++} codegen / compilation
    for NUTS), inference time, predictive log-likelihood (PLL) on
    hold-out data, expected sample size (ESS,
    higher is better, \cite{geyer2011introduction}) and the rank normalized
    $\widehat{R}$ diagnostic (Rhat, closer to $1$ is better, \cite{vehtari2020rank}).
  }\label{fig:blr_nschools}
  \begin{subfigure}[b]{\textwidth}
    \centering
    \caption{Bayesian logistic regression ($2000$ rows, $10$ features).
      Both LIC and Stan (NUTS) amortize the upfront compilation cost with accelerated inference times.
      LIC achieves comparable PLL to NUTS \cite{hoffman2014no} and other methods,
      and yields ESS comparable to NUTS and higher than any other method.
      The $\widehat{R}$ of LIC is close to $1$ (similar to NUTS and lower than
      all other methods).
    }\label{fig:blr}
    \includegraphics[width=\linewidth]{Figures/blr_pll_type1.pdf}
    \includegraphics[width=0.6\linewidth]{Figures/blr_ess_rhat_type1.pdf}
  \end{subfigure}
  \begin{subfigure}[b]{\textwidth}
    \centering
    \caption{n-schools (1000 schools, 8 states, 5 districts, 5 types).
      Again, increased compilation times are offset by accelerated inference times.
      In this case, LIC achieves comparable PLL to NUTS while simultaneously producing
      higher ESS and lower $\widehat{R}$, suggesting that the resulting posterior samples
      are less autocorrelated and provide a more accurate posterior approximation.
    }\label{fig:nschools}
    \includegraphics[width=\linewidth]{Figures/nschools_pll_type1.pdf}
    \includegraphics[width=0.66\linewidth]{Figures/nschools_ess_rhat_type1.pdf}
  \end{subfigure}
\end{figure*}

@@

% TODO: include in camera ready
% \subsubsection*{Acknowledgements}
% Feynman Liang is supported by a NPSC Graduate Fellowship.
% This work was completed during an internship at Facebook.
